---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ctasval

<!-- badges: start -->
[![R-CMD-check](https://github.com/IMPALA-Consortium/ctasval/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/IMPALA-Consortium/ctasval/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of ctasval is to provide structured code to validate {ctas} performance on study data.

## Installation

You can install the development version of ctasval from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("IMPALA-Consortium/ctas")
devtools::install_github("IMPALA-Consortium/ctasval")
```

## ctasval

`ctasval` adds three anomalous sites to the data set for each iteration and tries to detect them using {ctas}.
It samples from the study site pool to first determine the number of patients and then samples a sufficient
number of patients from the study patient pool.

```{r}

library(pharmaversesdtm)
library(dplyr)
library(future)
library(ctasval)
library(stringr)


set.seed(1)

df_prep <- prep_sdtm_lb(pharmaversesdtm::lb, pharmaversesdtm::dm, scramble = TRUE)

df_filt <- df_prep %>%
  filter(parameter_id %in% c("Alkaline Phosphatase", "Alanine Aminotransferase")) %>%
  filter(! grepl("UNSCH", VISIT))

plan(multisession, workers = 6)

ctas <- ctasval(
  df = df_filt,
  fun_anomaly = c(anomaly_average, anomaly_sd, anomaly_autocorr),
  feats = c("average", "sd", "autocorr"),
  parallel = TRUE,
  iter = 100,
  n_sites = 3,
  anomaly_degree = c(0, 0.1, 0.25, 0.5, 1, 2, 5),
  thresh = 1
)

plan(sequential)

ctas
```

### Performance Metrics

```{r}
ctas$result %>%
  knitr::kable(digits = 3)
```

### Anamolous Sites

Anomalous Sites and their scores can be reviewed.

```{r}
ctas$anomaly %>%
  select(
    iter,
    anomaly_degree,
    feats,
    parameter_id,
    site,
    subject_id,
    timepoint_rank,
    result,
    score
  ) %>%
  arrange(iter, desc(anomaly_degree), parameter_id, feats, site, subject_id, timepoint_rank) %>%
  head(50) %>%
  knitr::kable()

```
