---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ctasval

<!-- badges: start -->

[![R-CMD-check](https://github.com/IMPALA-Consortium/ctasval/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/IMPALA-Consortium/ctasval/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

The goal of ctasval is to provide structured code to validate {ctas} performance on study data.

## Installation

You can install the development version of ctasval from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("IMPALA-Consortium/ctas")
devtools::install_github("IMPALA-Consortium/ctasval")
```

## ctasval

`ctasval` adds three anomalous sites to the data set for each iteration and tries to detect them using {ctas}. It samples from the study site pool to first determine the number of patients and then samples a sufficient number of patients from the study patient pool.

```{r}

library(pharmaversesdtm)
library(dplyr)
library(future)
library(ctasval)
library(ggplot2)
library(tidyr)
library(arrow)



set.seed(1)

df_prep <- prep_sdtm(
  lb = pharmaversesdtm::lb,
  vs = pharmaversesdtm::vs,
  dm = pharmaversesdtm::dm,
  scramble = TRUE
)

df_filt <- df_prep %>%
  filter(parameter_id %in% c("Pulse Rate", "Alanine Aminotransferase")) %>%
  filter(! grepl("UNSCH", timepoint_1_name) & !timepoint_1_name %in% c("AMBUL ECG REMOVAL", "RETRIEVAL"))

```

**Note:** On some systems, running multiple ctasval chunks with multiprocessing in the same R session may cause hangs or memory issues. If you experience this, restart your R session between running different ctasval chunks.

```{r eval=FALSE}
plan(multisession, workers = 6)

ctas <- ctasval(
  df = df_filt,
  fun_anomaly = c(
    anomaly_average,
    anomaly_sd,
    anomaly_autocorr2,
    anomaly_lof,
    anomaly_range,
    anomaly_unique_value_count_relative
  ),
  feats = c("average", "sd", "autocorr", "lof", "range", "unique_value_count_relative"),
  parallel = TRUE,
  iter = 50,
  n_sites = 3,
  anomaly_degree = c(0, 0.25, 0.5, 0.75, 1, 5, 10),
  thresh = 1.3,
  site_scoring_method = "ks"
)

arrow::write_parquet(ctas$result, "ctas_ks_result.parquet")

ctas$anomaly %>%
  select(- fun_anomaly) %>%
  arrow::write_parquet("ctas_ks_anomaly.parquet")
```

### Performance Metrics

```{r}
ctas_result <- arrow::read_parquet("ctas_ks_result.parquet")

ctas_result %>%
  tidyr::pivot_longer(c(tpr, fpr), values_to = "metric", names_to = "metric_type") %>%
  ggplot(aes(log(anomaly_degree),  metric)) +
  geom_line(aes(linetype = metric_type)) +
  geom_point() +
  facet_grid(parameter_id ~ feats) +
  theme(legend.position = "bottom")
```


### Mixed Effect Scoring

```{r eval = FALSE}
plan(multisession, workers = 6)

ctas <- ctasval(
  df = df_filt,
  fun_anomaly = c(
    anomaly_average,
    anomaly_sd,
    anomaly_autocorr2,
    anomaly_lof,
    anomaly_range,
    anomaly_unique_value_count_relative
  ),
  feats = c("average", "sd", "autocorr", "lof", "range", "unique_value_count_relative"),
  parallel = TRUE,
  iter = 50,
  n_sites = 3,
  anomaly_degree = c(0, 0.25, 0.5, 0.75, 1, 5, 10),
  thresh = 1.3,
  site_scoring_method = "mixedeffects"
)

arrow::write_parquet(ctas$result, "ctas_mixedeffects_result.parquet")

ctas$anomaly %>%
  select(- fun_anomaly) %>%
  arrow::write_parquet("ctas_mixedeffects_anomaly.parquet")

```

```{r}
ctas_result <- arrow::read_parquet("ctas_mixedeffects_result.parquet")

ctas_result %>%
  tidyr::pivot_longer(c(tpr, fpr), values_to = "metric", names_to = "metric_type") %>%
  ggplot(aes(log(anomaly_degree),  metric)) +
  geom_line(aes(linetype = metric_type)) +
  geom_point() +
  facet_grid(parameter_id ~ feats) +
  theme(legend.position = "bottom")

```

### Average Box Plot Scoring

```{r eval=FALSE}
plan(multisession, workers = 6)

ctas <- ctasval(
  df = df_filt,
  fun_anomaly = c(
    anomaly_average,
    anomaly_sd,
    anomaly_autocorr2,
    anomaly_lof,
    anomaly_range,
    anomaly_unique_value_count_relative
  ),
  feats = c("average", "sd", "autocorr", "lof", "range", "unique_value_count_relative"),
  parallel = TRUE,
  iter = 50,
  n_sites = 3,
  anomaly_degree = c(0, 0.25, 0.5, 0.75, 1, 5, 10),
  site_scoring_method = "avg_feat_value"
)

arrow::write_parquet(ctas$result, "ctas_avgfeat_result.parquet")

ctas$anomaly %>%
  select(- fun_anomaly) %>%
  arrow::write_parquet("ctas_avgfeat_anomaly.parquet")

```

```{r}
ctas_result <- arrow::read_parquet("ctas_avgfeat_result.parquet")

ctas_result %>%
  tidyr::pivot_longer(c(tpr, fpr), values_to = "metric", names_to = "metric_type") %>%
  ggplot(aes(log(anomaly_degree),  metric)) +
  geom_line(aes(linetype = metric_type)) +
  geom_point() +
  facet_grid(parameter_id ~ feats) +
  theme(legend.position = "bottom")

```


### Anamolous Sites

Anomalous Sites and their scores can be reviewed.

```{r}
ctas_anomaly <- arrow::read_parquet("ctas_avgfeat_anomaly.parquet")

ctas_anomaly %>%
  select(
    iter,
    anomaly_degree,
    feats,
    parameter_id,
    site,
    subject_id,
    timepoint_rank,
    result,
    score
  ) %>%
  arrange(iter, desc(anomaly_degree), parameter_id, feats, site, subject_id, timepoint_rank) %>%
  head(50) %>%
  knitr::kable()

```

```{r}
plan(sequential)
```


## Publication

these commands render publication to `pub/`

to edit publication edit `inst/pub/ctasvalidation.qmd`

```{r eval=FALSE}
quarto::quarto_render("inst/pub/ctasvalidation.qmd", output_format = "html")
quarto::quarto_render("inst/pub/ctasvalidation.qmd", output_format = "docx")
quarto::quarto_render("inst/pub/ctasvalidation.qmd", output_format = "pdf")
quarto::quarto_render("inst/pub/ctasvalidation.qmd", output_format = "latex")

```

