---
title: "autocorr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{autocorr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ctasval)
library(ggplot2)
library(tidyr)
```

# Comparing Different Approaches for Introducing Autocorrelation to Clinical Measurements

- `anomaly_autocorr` uses a sinus function
- `anomaly_autocorr2` always adds a fraction of the lag value to the current value

```{r}
set.seed(123)
df_prep <- prep_sdtm_lb(pharmaversesdtm::lb, pharmaversesdtm::dm, scramble = TRUE)

df_filt <- df_prep %>%
 filter(parameter_id == "Alkaline Phosphatase")

# with this wrapper functions will always generate an anomalous site with the same patients
set_seed <- function(fun, seed = 1){
  fun_seeded <- function(...) {
    set.seed(seed)
    fun(...)
  }
}


plot_anomaly <- function(df, fun, anomaly_degree, seed) {
  fun_seeded <- set_seed(fun, seed)
  df_anomaly <- fun_seeded(df, anomaly_degree = anomaly_degree)
  
  ggplot(df, aes(x = timepoint_rank, y = result, group = subject_id)) +
    geom_line(color = "black") +
    geom_line(data = df_anomaly, color = "tomato") +
    coord_cartesian(xlim = c(0, max(df_anomaly$timepoint_rank)))
}

plot_anomaly(df_filt, anomaly_autocorr, anomaly_degree = 0.7, seed = 3)
plot_anomaly(df_filt, anomaly_autocorr2, anomaly_degree = 0.7, seed = 3)

```

# Compare Autocorrelation Values

```{r}
get_autocorr <- function(df, fun, anomaly_degree) {
  fun_seeded <- set_seed(fun)
  df_anomaly <- fun_seeded(df, anomaly_degree = anomaly_degree)
  
  df_anomaly %>%
    filter(! is.na(result)) %>%
    summarise(
      autocorr = ctas:::calculate_autocorrelation(result),
      n = n(),
      .by = subject_id
    ) %>%
    mutate(
      autocorr = ifelse(n <= 2, 0, autocorr)
    )
}

get_autocorr(df_filt, anomaly_autocorr, anomaly_degree = 0.7)
```

```{r}
df_grid <- tibble(
    anomaly_degree = list(seq(0, 1, 0.1)),
    fun = list(anomaly_autocorr, anomaly_autocorr2),
    fun_name = c("anomaly_autocorr", "anomaly_autocorr2")
  ) %>%
  unnest(anomaly_degree)

df_autocorr <- df_grid %>%
  mutate(
    autocorr = purrr::map2(fun, anomaly_degree, ~ get_autocorr(df_filt, .x, .y))
  ) %>%
  unnest(autocorr)
```

```{r fig.height=5, fig.width=7}
df_autocorr %>%
  ggplot(aes(as.factor(anomaly_degree), autocorr, fill = fun_name)) +
  geom_boxplot() +
  theme(legend.position = "bottom")
```

# Conclusion

`anomaly_autocorr2` which adds a fraction of the previous values increases the autocorrelation more than `anomaly_autocorr` which uses a sinus function. 

`ctas:::calculate_autocorrelation` only uses lag 1. To detect autocorrelation induced by the sinus function other lags would need to be considered.

## Autocorrelation of a Sinus Function considering lag 1 to 9

```{r}
acf(sin(seq(1:10)))
```

## Autocorrelation of a random patient for Alkaline Phosphatase measurements considering lag 1 to 9

```{r}
df_filt %>%
  filter(dense_rank(subject_id) == 1) %>%
  pull(result) %>%
  acf
```



